Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-13 paper.tex
new bib item first section about anthropomorphism definitions

In this section we will clarify the terminology that we use as we have found conflicting understandings of the term ``anthropomorphism'' in the related literature. We first present anthropomorphism as a social phenomenon. Then we propose to use another term, namely ``anthropomorphic form'', to refer to the human-inspired design of artifacts. Finally, we focus on anthropomorphism in Human-Robot Interaction, as it naturally brings together (and needs to question) the usage of an anthropomorphic design of robots and its consequences for the interaction with the human user.

@@

Anthropomorphism is an interesting phenomenon because it tells us something about the relationship between the human who anthropomorphizes, and the non-human agent which is anthropomorphized. Though commonly observed, it appears controversial, at the same time, that humans tend to perceive human-like physical appearance (\eg faces), emotional states, or inner mental states and motivations in non-human agents such as animals, natural forces, religious agents, technological gadgets, or mechanical devices \citep{epley_when_2008}.\footnote{According to \cite{epley_when_2008}, a non-human agent can be anything that acts -- or is believed to act -- with apparent independence.}  Anthropomorphism has received remarkable attention across several domains, such as natural sciences, psychology, Human-Computer Interaction, and Human-Robot Interaction. However, the term is used in different senses throughout these disciplines \citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings.

Originally, \textbf{anthropomorphism} comes from the Greek \textit{`anthropos'} for `man' (or `human') and \textit{`morphe'} for `form / structure' (or `shape'). It can generally be understood as people's tendency to perceive a non-human agent as if it had some human characteristics. In turn, how a person perceives an agent further determines how she relates to and behaves towards the agent, or how she may behave in light of this agent \citep{epley_when_2008}. By this, anthropomorphism can be seen as a special kind of relationship between the human and the non-human agent. This relationship is often characterized by the fact that the human attributes or ascribes uniquely human characteristics to the non-human entity or artifact. The most prominently anthropomorphized non-human agents are probably animals and robots \citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism there exists also \textbf{animism}, which can be understood as the \textit{"attribution of life to the nonliving"} \hl{in Schmitz 2011, cite Piaget original text!}. Using this definition, animism differs in two main aspects from anthropomorphism:
First, the property which is ascribed differs: it is not human-unique but applies to all living beings (life-likeness). For instance, attributing life-likeness to a stone must not necessarily be labeled as anthropomorphism, since being alive is not a human unique property but applies to all living beings.
The second difference comes from the instance (or better things) to which life is attributed: all living entities are excluded in this definition. In other words, pets as non-human but living beings can only be anthropomorphized (because they are not human) but not animized (because they are already alive).
Often synonymously used to animism is the term \textbf{zoomorphism}, which applies to the cases when non-lifelike instances / things are associated with animalistic attributes, excluding human-specific traits. Both anthropomorphism and zoomorphism can be embraced by the concept of \textbf{life-likeness}. \footnote{The inverse process would be called dehumanization \cite{haslam_dehumanization:_2006}, or mechanomorphism \cite{caporael_anthropomorphism_1986}.}

Commonly, anthropomorphism can also be understood as \textbf{humanization} or
\textbf{personification} of anything other than a human being. The phenomenon of
personifying something has ancient roots, \textit{e.g.} in mythology but also in
storytelling. In prehistoric artworks and illustrations, zoomorphic
(animal-shaped) or anthropomorphic shapes were already used to represent natural
forces or great spirits, for instance. A more recent example is Milo Winter's
illustration from 1919, \textit{"The North Wind and the Sun"} (Figure
\ref{fig:1}), in which both north wind and sun are given a human face
(anthropomorphized). Similarly, when parents explain their children about
nature, they tend to use anthropomorphisms, \textit{e.g.} saying the skies would
cry when it rains. More generally, an explanation for why people
anthropomorphize is that by doing so, the subject is made \emph{graspable},
\emph{understandable}, \emph{predictable}, and one feel \emph{more empathetic}
toward it\fxnote{important part. Emphasize this 4 facets, eg by saying we build
on top of them}. For instance, people commonly anthropomorphize their pets,
ascribing mental states, intentions or feelings to them
\cite{eddy_attribution_1993}. Interestingly, pet-ownership seems to have a
significant impact on human-robot interaction: several studies suggest that
pet-owners are more likely to anthropomorphize technologies and robots than non
pet-owners \hl{references!}.

@@

@@

\fxfatal{inconsistency: we already give explanations for anthropomorphism in previous paragraph...}

Anthropomorphism represents just one of many examples of induction whereby
``people reason about an unknown stimulus based on a better-known representation
of a related stimulus"~\cite{epley_when_2008}, in this case reasoning about a
non-human agent based on representation of the self or other humans.

According to Lee \etal~\cite{lee_human_2005}, there are two main perspectives in
explaining people's tendency to anthropomorphize. First one explains
anthropomorphism from the design of the artifact. It assumes that humans
directly respond to life-like or social cues that an object or system emits,
without thoughtful mental processing, by simply applying stereotypes and
heuristics to it. In fact, from early childhood on, humans are inherently
well-trained to perceive life \cite{epley_seeing_2007}. Schmitz
\cite{schmitz_concepts_2011} describes that within the visual scope of design,
the outer appearance can have an important impact on the overall perception of
an object. The basic assumption here is that if an artifact appears much like a
human, it is likely to be treated similar to a human. If this explanation of
anthropomorphism is correct, people may respond automatically to social cues
emitted by a robot, and apply human-human social schemas and norms to these
interactions.

The second perspective applies a human-centered, cognitive viewpoint where
anthropomorphism is described through people's specific mental model they
construct about how an artifact works the way it does.  We then
anthropomorphize because it allows us to explain things we do not understand in
terms that we do understand, and what we understand best is ourselves as human
beings. This is consistent with the \emph{familiarity
thesis}~\cite{hegel_understanding_2008} which claims that we understand the
world based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in
their environment and make inferences about it based on what is familiar to
them -- humans and human behavior, for instance.  This point of view implicitly
builds on a person's ability to attribute mental states to oneself and others
(\ie the availability of a \emph{theory of mind}~\cite{premack1978does} -- the
link between a tendency to anthropomorphize and the engagement in the
attribution of mental states to other humans has been recently demonstrated at
the brain level in~\cite{cullen2013individual}). A theory of mind for other
agents enables us to attribute intentionality to those agents
\cite{leslie_pretense_1987,admoni_multi-category_2012}. Previous research
examined the validity of the mental model concept with various kinds of robots
\cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models of anthropomorphic robots in contrast to
mechanic ones \cite{kiesler_mental_2002}.

According to Lee \textit{et al.} \cite{lee_human_2005}, there are two main
perspectives in explaining people's tendency to anthropomorphize. The first
one explains anthropomorphism from the design of the artifact
(anthropomorphic form in the design). It assumes that humans directly
respond to life-like or social cues that an object or system emits, without
thoughtful mental processing, by simply applying stereotypes and heuristics
to it. In fact, from early childhood on, humans are inherently well-trained
to perceive life \cite{epley_seeing_2007}. Schmitz
\cite{schmitz_concepts_2011} describes that within the visual scope of
design, the outer appearance can have an important impact on the overall
perception of an object. The basic assumption here is that if an artifact
appears much like a human, it is likely to be treated similar to a human. If
this explanation of anthropomorphism is correct, people may respond
automatically to social cues emitted by a robot, and apply human-human
social schemas and norms to these interactions.

The second perspective applies a human-centered, cognitive viewpoint where
anthropomorphism is described through people's specific mental model they
construct about how an artifact works the way it does. According to v.
Foerster \hl{ref} \textit{"we anthropomorphize because it allows us to
explain things we do not understand in terms that we do understand, and what
we understand best is ourselves as human beings"}
\cite{hegel_understanding_2008}. This is consistent with the \hl{familiarity
thesis (18)} \cite{hegel_understanding_2008} which claims we understand the
world based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in
their environment and make inferences about it based on what is familiar to
them. This point of view implicitly builds on a person's ability to attribute
mental states to oneself and others, which is called \textit{Theory of
Mind}~\cite{Premack1978}. A theory of  mind for other agents enables us to
attribute intentionality to those agents
\cite{leslie_pretense_1987,admoni_multi-category_2012}. If a system behaves much
like a human being (\eg emits a human voice), people's mental model of the
system's behavior may approach their mental model of humans, though the model
may differ in some important aspects \hl{ref? Schmitz?}. In turn, people's
estimation of a robot's knowledge model and its capabilities affects the way
they relate to it. Previous research examined the validity of the mental model
concept with various kinds of robots
\cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models about anthropomorphic robots in contrast to
mechanic ones \cite{kiesler_mental_2002}. Later in this article, we may however
come to question this result for longer human-robot interactions, after one
get's acquantained with the robot's behaviour.

@@

@@

However, not everything colloquially labeled as \emph{anthropomorphism} really is anthropomorphism. Different, even partially conflicting understandings of anthropomorphism exist across and within disciplines \cite{duffy_anthropomorphism_2002}. \cite{epley_when_2008} try to narrow down and clarify the understanding of anthropomorphism in psychology by explaining what it is \textit{not}. The authors list four things
that anthropomorphism does not include: First, anthropomorphism does not include
behavioral descriptions of observable actions. Anthropomorphism requires going
beyond what is directly observable. Second, anthropomorphism does not merely
entail animism, as animate life is not a uniquely human-like characteristic.
Third, anthropomorphism does not include any requirement of reasoned or
reflective endorsement of an inference (\hl{the strength of anthropomorphic
inferences can vary} \footnote{for further explanation the reader may refer to
the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not
necessarily inaccurate. A psychological theory should take this (in)accuracy in
anthropomorphism into account and predict variability in the tendency to
perceive human-like traits in non-human agents.  In conclusion, it has been
noted that anthropomorphism is a very complex phenomenon - in itself, its
understanding, and when studying it. A common formal ground is missing; however,
it is important to be clear about what exactly we mean when using the term
anthropomorphism.

>>>>

In this section we will try to create a common ground in the understanding of anthropomorphism since we have found conflicting definitions of the term. We first present anthropomorphism as a social phenomenon. Then we propose to use another term, namely ``anthropomorphic form'', to refer to the human-inspired design of artifacts. Finally, we focus on anthropomorphism in Human-Robot Interaction, as it naturally brings together (and needs to question) the usage of an anthropomorphic design of robots and its consequences for the interaction with the human user.

@@

Anthropomorphism is an interesting phenomenon because it tells us something about the relationship between the human who anthropomorphizes, and the non-human agent which is anthropomorphized. Though commonly observed, it appears controversial, at the same time, that humans tend to perceive human-like physical appearance (\eg faces), emotional states, or inner mental states and motivations in non-human agents such as animals, natural forces, religious agents, technological gadgets, or mechanical devices \citep{epley_when_2008}.\footnote{According to \cite{epley_when_2008}, a non-human agent can be anything that acts -- or is believed to act -- with apparent independence.}  Anthropomorphism is one of not so many other topics that has received remarkable attention across various different domains, such as natural sciences, developmental and social psychology, Human-Computer Interaction (HCI), and Human-Robot Interaction (HRI). However, the term is used in different senses throughout these disciplines \citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings.

Originally, the term \textbf{anthropomorphism} comes from the Greek \textit{`anthropos'} for `man' (or `human') and \textit{`morphe'} for `form / structure' (or `shape'). Anthropomorphism is generally understood as people's tendency to perceive a non-human agent as if it had some human characteristics or traits. In turn, how a person perceives an agent further determines how she relates to and behaves towards the agent, or how she may behave in light of this agent \citep{epley_when_2008}.
As such, anthropomorphism can be seen as a special kind of relationship between a human interacting with a non-human agent. This anthropomorphic relationship is often characterized by the fact that the human attributes or ascribes human-unique characteristics to the non-human agent. Since anthropomorphism becomes observable through these attributions, some definitions claim that anthropomorphism is not the perception but rather the attribution of human attributes to non-human agents, and name thoughts, emotions, and intentions as examples. The most prominently anthropomorphized non-human agents are probably animals and robots \citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism there also exists the term \textbf{animism}, which can be defined as the attribution of life to the nonliving.\footnote{The study of animism was pioneered by \hl{Piaget} in the 1920s with his work on child's conception of the  world and of the living. He found that children tend to regard many inanimate objects as capable of sensations, emotions and intentions, and called this view `animism'. Since then, children's inferences about the world have been studied intensively and it has been found that especially younger children (according to \hl{Piaget} up to the age of about 10 years) tend to interpret physical phenomena in terms of intention on the part of the inanimate subject, for instance \textit{``the sun is hot because it wants to make people warm''}. However, there is an agreement that children's views and concepts of living, animal, and plant evolve as they grow up.} Using this definition, animism differs in two main aspects from anthropomorphism:
First, the property which is ascribed differs: life-likeness is not human-unique but applies to all living beings. The second difference between anthropomorphism and animism stems from the agent (or better artifact) to which life is attributed. Living beings (most importantly animals) are naturally excluded from being animized. Consequently, while animals can be anthropomorphized (perceived to have human unique characteristics), they cannot be animized as they are already living beings. (Similarly to the fact that humans cannot be anthropomorphized as they are already humans.)
Often synonymously used to animism is the term \textbf{zoomorphism}, which applies to the cases when non-lifelike instances / things are associated with animalistic attributes, excluding human-specific traits. Both anthropomorphism and animism / zoomorphism can be embraced by the concept of \textbf{life-likeness}.

There exist also inverse phenomenons to anthropomorphism, which can be found in the literature as \textbf{dehumanization} \citep{haslam_dehumanization:_2006}, and \textbf{mechanomorphism} \citep{caporael_anthropomorphism_1986}. However, we do not treat these phenomenons here.

@@

To illustrate better how anthropomorphism differs from attributions of animacy or life-likeness, it is worth looking at paraphrases and examples of it. That is to say, anthropomorphism can alternatively be understood as \textbf{humanization} or \textbf{personification} of anything other than a human being. Personification has ancient roots, \eg in mythology but also in storytelling. Anthropomorphism was also used to describe and be able to predict natural forces or anticipate great spirits, for instance. An illustration of an anthropomorphism related to nature is shown in Figure \ref{fig:1}. In this Milo Winter (1919) drawing \textit{"The North Wind and the Sun"}, both north wind and sun are given a human face.

@@

But why do people anthropomorphize? One general reason is that by anthropomorphizing the non-human subject, it is made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one can feel \emph{more empathetic} toward it, and possible anticipate the subject's behavior. \fxnote{important part. Emphasize this 4 facets, eg by saying we build on top of them} For instance, people commonly anthropomorphize their pets, ascribing mental states, intentions or feelings to them\footnote{Interestingly, pet-ownership seems to have a significant impact on human-robot interaction: several studies suggest that pet-owners are more likely to anthropomorphize technologies and robots than non pet-owners} \cite{eddy_attribution_1993}. But what are the motivations behind this, generally speaking?

Anthropomorphism represents just one of many examples of induction whereby ``people reason about an unknown stimulus based on a better-known representation of a related stimulus"~\cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans. \hl{Julia: this needs to go somewhere else ...}

According to \cite{lee_human_2005}, there are two main perspectives in explaining people's tendency to anthropomorphize.

First one explains anthropomorphism from ascertainable parts / the design of the non-human agent. This perspective assumes that humans directly respond to life-like or social cues that an object or system emits, without thoughtful mental processing, by simply applying stereotypes and heuristics to it. In fact, from early childhood on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}. \cite{schmitz_concepts_2011} describes that within the visual scope of design, the outer appearance of an object can have an important impact on how it is overall perceived. The basic assumption here is that if an artifact somehow resembles a human, it is likely to be treated similar to a human. If this explanation of anthropomorphism is correct, people may respond automatically to social cues emitted by an artificial embodied agent such as a robot, and apply human-human social schemas and norms to the occurring interactions.

\hl{Julia: change the order of the two explanations! that makes more sense in terms of previous paragraph and anth form as following paragraph}

The second perspective applies a human-centered, cognitive viewpoint where anthropomorphism is described through people's specific mental model they construct about how an artifact works the way it does. We then anthropomorphize because it allows us to explain things we do not understand in terms that we do understand, and what we understand best is ourselves as human beings. This is consistent with the \emph{familiarity thesis}~\cite{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in their environment and make inferences about it based on what is familiar to them -- humans and human behavior, for instance. This point of view implicitly builds on a person's ability to attribute mental states to oneself and others, which is called \textit{Theory of Mind}~\cite{Premack1978}. The link between a tendency to anthropomorphize and the engagement in the attribution of mental states to other humans has been recently demonstrated at the brain level in~\cite{cullen2013individual}. A theory of mind for other agents enables us to attribute intentionality to those agents \cite{leslie_pretense_1987,admoni_multi-category_2012}. If a system behaves much like a human being (\eg emits a human voice), people's mental model of the system's behavior may approach their mental model of humans, though the model may differ in some important aspects. In turn, people's estimation of a robot's knowledge model and its capabilities affects the way they relate to it. Previous research examined the validity of the mental model concept with various kinds of robots \cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \cite{kiesler_mental_2002}. Later in this article, we may however come to question this result for longer human-robot interactions, after one get's acquainted with the robot's behavior.

@@

Anthropomorphic forms based on  have already in prehistoric artworks and illustrations, anthropomorphic (but sometimes also animal-shaped) characters

@@

However, not everything colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.\footnote{For instance, it is not always clear whether a person (may it be an adult or a child) who gives and anthropomorphic response, believes that the other agent really thinks or acts like a human or whether she is using anthropomorphism in a metaphoric sense.\citep{leeds_childrens_1992}} Different, even partially conflicting understandings of anthropomorphism exist across and within disciplines \cite{duffy_anthropomorphism_2002}. \cite{epley_when_2008} try to narrow down and clarify the understanding of anthropomorphism in psychology by explaining what it is \textit{not}. The authors list four things
that anthropomorphism does not include: First, anthropomorphism does not include
behavioral descriptions of observable actions. Anthropomorphism requires going
beyond what is directly observable. Second, anthropomorphism does not merely
entail animism, as animate life is not a uniquely human-like characteristic.
Third, anthropomorphism does not include any requirement of reasoned or
reflective endorsement of an inference (\hl{the strength of anthropomorphic
inferences can vary} \footnote{for further explanation the reader may refer to
the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not
necessarily inaccurate. A psychological theory should take this (in)accuracy in
anthropomorphism into account and predict variability in the tendency to
perceive human-like traits in non-human agents.  In conclusion, it has been
noted that anthropomorphism is a very complex phenomenon - in itself, its
understanding, and when studying it. A common formal ground is missing; however,
it is important to be clear about what exactly we mean when using the term
anthropomorphism.
